We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-21 paper/paper.tex
edit

Later, \cite{2007MNRAS.374.1125M} used simulations to confirm the low
chances ($<0.5\%$) found in \cite{2005A&A...431..517K}.
This result was challenged by \cite{2009MNRAS.399..550L} who continue
to report high chances to find a planar configuration;
two more recent numerical experiment with high resolution simulations
(\cite{2013MNRAS.429..725S} and \cite{2016MNRAS.457.1931S}) report
contradicting results (low and high chances to find the observational
result, respectively) but precise figures are not quoted in either of
these three reports. 
Meanwhile, \cite{2013MNRAS.429.1502W} and \cite{2014ApJ...789L..24P} reported
on new tests using high resolution simulations and cosmological
volumes arriving at a chance of $6\%$ and $0.77\%$, respectively, to
have a satellite distribution as triaxial as the observations.

In the case of M31, observational studies have found a planar satellite
distribution in a special subset of $15$ satellites out of the 
total poulation of $27$ satellites \citep{2013ApJ...766..120C,
2013Natur.493...62I}, although its brightest satellites
do not show any special feature.
All published studies agree on the point that the spatial distribution of
the 15 to 27 brightest satellites are consistent with 
spherically symmetric distribution and easy to reproduce in LCDM
simulations \citep{2006AJ....131.1405K,
  2007MNRAS.374.1125M, 2013ApJ...766..120C}.    
Table \ref{table:M31} and Table \ref{table:MW} summarize some more
details on the results we have just mentioned \footnote{
Those tables also include the results from this paper using the
methodolgy we describe in the upcoming sections.}.

@@

Some of the difficulty in trying to reconcile and understand the
seemly conflicting or inconclusive results on the MW has its origin on
the frequentist fashion generally used to compute probabilities.
Usually, this process starts by building a high level parent sample in the
simulation and then counting how many elements has a subset
meeting some criteria. This has two incovenients.
The first is that the probability estimate is made against whatever
turns out to be typical in each simulation. 
A fair comparison across simulations would require first
characterizing all the simulations at the high level parent samples,
something that is difficult to do in practice. 
The second inconvenient is that it is hard to find systems that fully resemble
the LG while having enough resolution to say something about the
satellites, this means that the high level parent sample has a small
size, making it hard to derive robust probabilities by counting.

@@

The rest of the paper is structured as follows. 
In Section \ref{sec:DataSamples} we list the sources of the observational and
simulated data to be used throughout the paper.
In Section \ref{sec:SpatialMeasurements} we describe the methods we use to build halo pairs and 
characterize its satellite distributions.
In Section \ref{sec:results} we present our results to finally
conclude in Section \ref{sec:conclusions}.

>>>>

Later, \cite{2007MNRAS.374.1125M} used simulations to confirm the low
chances ($<0.5\%$) found in \cite{2005A&A...431..517K}.
This result was challenged by \cite{2009MNRAS.399..550L} who continued
to report high chances to find a planar configuration;
two more recent numerical experiment with high resolution simulations
(\cite{2013MNRAS.429..725S} and \cite{2016MNRAS.457.1931S}) reported
contradicting results (low and high chances to find the observational
result, respectively) but precise figures are not quoted in either of
these three reports. 
Meanwhile, \cite{2013MNRAS.429.1502W} and \cite{2014ApJ...789L..24P} published
the results new tests using high resolution simulations and cosmological
volumes arriving at a chance of $6\%$ and $0.77\%$, respectively, to
have a satellite distribution as triaxial as the observations.

In the case of M31, observational studies have found a planar satellite
distribution in a special subset of $15$ satellites out of the 
total poulation of $27$ satellites \citep{2013ApJ...766..120C,
2013Natur.493...62I}, although considering the satellites ranked by
luminosity does not show any special feature.
All published studies agree on the point that the spatial distribution of
the 15 to 27 brightest satellites are consistent with  spherically
symmetric distribution and easy to reproduce in LCDM simulations
\citep{2006AJ....131.1405K,2007MNRAS.374.1125M, 2013ApJ...766..120C}.    
Table \ref{table:M31} and Table \ref{table:MW} summarize some more
details on the results we have just mentioned \footnote{
Those tables also include the results from this paper using the
methodolgy we describe in the upcoming sections.}.

@@

Some of the difficulty in trying to reconcile and understand the
seemly conflicting or inconclusive results on the MW has its origin on
the frequentist fashion generally used to compute probabilities.
Usually, this process starts by building a high level parent sample in the
simulation and then counting how many elements has a subset
meeting some criteria. This has two incovenients.
The first is that the probability estimate is made against whatever
turns out to be typical in each simulation. 
A fair comparison across simulations would require first
characterizing all the simulations at the high level parent samples,
something that is difficult to do in practice. 
The second inconvenient is that the systems that fully resemble
the LG (i.e. its stellar mass content, morphology or kinematics) have
a low cosmological number density, this means that for the current
cosmological volumes in simulations the high level parent sample has a
small size, making it hard to derive robust probabilities by counting.

@@

The rest of the paper describes in detail our implementation and
results. It is structured as follows. 
In Section \ref{sec:DataSamples} we list the sources of the observational and
simulated data to be used throughout the paper.
In Section \ref{sec:SpatialMeasurements} we describe the methods we use to build halo pairs and 
characterize its satellite distributions.
In Section \ref{sec:results} we present our results to finally
conclude in Section \ref{sec:conclusions}.
